\title{
  Adaptive Aggregation of Recommender Systems
}
\author{
  \AuthorName{Olav Bjørkøy (olavfrih@stud.ntnu.no)}
  \AuthorAffiliation{Department of Computer Sciences\\Norwegian University of Science and Technology\\Trondheim, Norway}
}

\Abstract{
  \begin{abstract}
Modern recommender systems combine multiple standard recommenders
in order to leverage disjoint patterns in available data.
These aggregations are done on a generalized level
by estimating weights that result in an optimal combination.
However, we posit these systems have an important weakness.
There exists an underlying, misplaced subjectivity to relevance prediction.
In an aggregation of different recommenders,
each chosen algorithm reflects one view of 
how each user and item \emph{should} be modeled.
We believe this selection should 
be adaptively and automatically chosen based on
how accurate each prediction is likely to be for each user and item.
This paper presents a novel method for prediction aggregation,
called \emph{adaptive recommenders}.
Multiple recommender systems are combined on a per-user and per-item basis
by estimating how accurate each recommender will be for the current user and item.
This is done by creating a set of secondary error estimating recommenders.
Our initial experiments are promising, showing that adaptive 
recommenders can outperform basic methods and simple aggregation approaches.
%As far as we know, this type of adaptive prediction aggregation
%has not been done before.
  \end{abstract}
}

\maketitle

