\section{Three Hypotheses}
\label{sec:hypotheses}

To evaluate whether or not \emph{stacked user modeling}
can mitigate the latent subjectivity problem,
we need a set of hypotheses.
To solve the subjectivity problem we need our modeling method
to adaptively aggregate a set of predictions based on the current user and item.
In other words, by automatically adapting how a set of disjoint recommenders
are combined, based on each user and item, we should be able to achieve a
result that is better than each of the stand-alone recommenders.
This adaptive method should also outperform other, generalized aggregation approaches.

This paper will consider three hypotheses (H1-H3) 
to establish whether adaptive aggregation through stacked user modeling is a viable technique.
H1 and H2 will consider the approach in regard to prediction aggregation
in a recommendation scenario. H3 will consider using this approach for
rank aggregation in an information retrieval scenario.
Let us start with prediction aggregation:

{
  \itshape
  $\mathbf{H1}$: The accuracy of relevance predictions can be improved
  by blending multiple modeling methods on a per-user and per-item basis.
}

It stands to reason that if a recommender system is indeed impaired
by the subjective selection of modeling methods,
an adaptive combination of these methods should outperform each of the individual approaches.
Second:

{
  \itshape
  $\mathbf{H2}$: An adaptive aggregation method can outperform global and generalized 
  blending methods.
}

If our assumption that model aggregation inherits the subjective nature of its chosen parts,
an adaptive aggregation without such misplaced subjectivity should outperform a
generalized combination.
Third:

{
  \itshape
  $\mathbf{H3}$: The result set from an information retrieval query
  can be personalized by stacked user modeling, where the retrieval scores are considered 
  standard input signals.
}

As described in Section \ref{subsec:signals},
modern search engines combines multiple ranking functions called signals into a final results list.
Each signal, be it an IR score, a metric like PageRank, or the result of a user modeling method,
contributes to the final item ranking.
We shall use H3 to see whether or not stacked user modeling can be used for this type of rank aggregation,
where a set of recommender systems constitute a set of input signals.

By answering our three hypotheses, it should become clear whether or not our approach
is a valid method of predicting the relevance of an item to a user.
These hypothesess will be tested in the next chapter, but first, let us build
the model we shall test.

