\section{Adaptive Prediction Aggregation}

\begin{table}[t]
  \begin{tabular*}{\textwidth}{ l p{3cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.5cm} }
    \toprule
      ~ & \emph{method} & 
      $d_1$ & $d_2$ & $d_3$ & $d_4$ & $d_5$ \\ 
    \midrule
    S & svd1          & 1.2389	  & 1.1260	  & 1.1327	  & 1.1045	  & 1.1184	 \\
    S & svd2          & 1.2630	  & 1.1416    & 1.1260	  & 1.1458	  & 1.1260	 \\
    S & svd3          & 1.0061	  & 0.9825	  & 0.9830	  & 0.9815	  & 0.9797	 \\
    S & svd4          & 1.0040	  & 0.9830	  & 0.9849	  & 0.9850	  & 0.9798	 \\
    S & slope\_one    & 1.1919	  & 1.0540	  & 1.0476	  & 1.0454	  & 1.0393   \\
    S & item\_avg     & 1.0713	  & 0.9692	  & 0.9662	  & 0.9683	  & 0.9725	 \\
    S & baseline       & 1.0698	  & 0.9557	  & 0.9527	  & 0.9415	  & 0.9492	 \\
    S & cosine   	    & 1.1101	  & 0.9463	  & 0.9412	  & 0.9413	  & 0.9382	 \\
    S & knn       	  & 1.4850	  & 1.1435	  & 1.1872    & 1.2156	  & 1.2022	 \\
    \midrule                                                                    
    A & median    	  & 0.9869	  & 0.8886	  & 0.8857    & 0.8857	  & 0.8855	 \\
    A & average    	  & 0.9900	  & 0.8536	  & 0.8525	  & 0.8525	  & 0.8519	 \\
    A & stacked       & \textbf{0.9324}	  & \textbf{0.8015}	  & \textbf{0.7993}  & \textbf{0.8238} & \textbf{0.8192} \\
    \bottomrule
  \end{tabular*}
  \caption[Results from Experiment 1]{
    Results from Experiment 1:
    Each cell gives an RMSE value for a method on a subset ($d_x$) of our dataset.
    As these are error measures, on the same scale as our predictions,
    lower values indicate better results. Bold values indicate the best result in each column.
    The first column refers to the type of predictions, either singular (S) or aggregation (A).}
  \label{table:results:e1}
\end{table}

Our first hypothesis, H1, states that:
{
  \itshape
  the accuracy of relevance predictions can be improved
  by blending multiple modeling methods on a per-user and per-item basis.
}
The second hypothesis, H2, states:
{
  \itshape
  an adaptive aggregation method can outperform global and generalized 
  blending methods.
}

In order to verify these hypotheses, we performed adaptive prediction aggregation
through stacked user modeling on the five datasets described in the previous section.
Table \ref{table:results:e1} gives the results from this experiment.
Each cell corresponds to the RMSE values for each dataset,
for each recommender and aggregation approach.
The bottom entry in this table refers to our stacked user modeling method.
As seen in this table, stacked user modeling achieved
lower RMSE values than any of the other applied methods.

Statistics from the results in Table \ref{table:results:e1} 
are given in Table \ref{table:results:e1:sum}.
These values are the minimum, maximum and mean values
for each of the methods. We also include
the standard deviation ($\sigma$) for each method,
across our datasets.
This table confirms the results from the full results table:
Our stacked user modeling approach improves the mean performance
of our system.
The mean performance, along with the standard deviation
are shown in Figure \ref{plot:rmse}.

\begin{table}[t]
  \begin{tabular*}{\textwidth}{ l p{3cm} p{2cm} p{2cm} p{2cm} p{2cm} }
    \toprule
      ~ & \emph{method} & 
      \emph{min} & \emph{max} & \emph{mean} & $\sigma$\\
    \midrule
    S & svd1          & 1.1045	& 1.2389	& 1.1441	& 0.2197 \\
    S & svd2          & 1.1260	& 1.2630	& 1.1605	& 0.2277 \\
    S & svd3          & 0.9797	& 1.0061	& 0.9865	& 0.0991 \\
    S & svd4          & 0.9798	& 1.0040	& 0.9873	& \textbf{0.0924} \\
    S & slope\_one    & 1.0393	& 1.1919	& 1.0756	& 0.2415 \\
    S & item\_avg     & 0.9662	& 1.0713	& 0.9895	& 0.2023 \\
    S & baseline       & 0.9415	& 1.0698	& 0.9738	& 0.2196 \\
    S & cosine   	    & 0.9382	& 1.1101	& 0.9754	& 0.2595 \\
    S & knn       	  & 1.1435	& 1.4850	& 1.2467	& 0.3487 \\
    \midrule            
    A & median    	  & 0.8855	& 0.9865	& 0.9065	& 0.2005 \\
    A & average    	  & 0.8519	& 0.9900	& 0.8801	& 0.2344 \\
    A & stacked       & \textbf{0.7993}	& \textbf{0.9324}	& \textbf{0.8352}	& 0.2225 \\
    \bottomrule
  \end{tabular*}
  \caption[Statistics from Experiment 1]{
      Statistics from Experiment 1: 
      This table shows the minimum, maximum and mean RMSE for each of the methods in Table \ref{table:results:e1}.
      The last column shows the standard deviation ($\sigma$) of each method.
      As before, lower values indicate better results, and bold values
      represent the best result in each column.
    }
  \label{table:results:e1:sum}
\end{table}

\input{plot.rmse}

\input{plot.datasets}

Let us take a look at the standard deviation measures from the different methods.
As seen in Figure \ref{plot:rmse}, 
most of the methods, including the stacked models,
exhibit quite a lot of variation in their results.
If these variations occured as a result of unstable
predictions of the same dataset, this would be a substantial problem,
resulting in unreliable predictions.
However, as seen in Figure \ref{plot:datasets},
the standard deviation is mostly caused by the differing
performance across the varying datasets.
As we see, the performance of each of the aggregation methods,
as well as the best performing standard recommender,
follow each other closely. At the same time,
performance varies across the different datasets,
which results in high values for $\sigma$.

What does this mean for hypotheses H1 and H2?
Expressed in terms of this experiment,
H1 posits that stacked user modeling should outperform each of the standard modeling methods
in Table \ref{table:results:e1}.
The adaptive methods blend the results of multiple predictors by estimating the accuracy
on a per-item and per-user basis, satisfying the formulation of H1.

By outperform we mean that our model should have a lower
mean RMSE score than the other singular methods. As we can see in Table \ref{table:results:e1:sum},
\emph{H1 is confirmed for these methods and this dataset}.
While we can not generalize too much on this basis, 
the fact that this dataset is a common testing ground for recommender systems,
that RMSE is the de facto measure for determining performance,
and because of our 5-fold cross-validation, the results allow us 
to confirm hypothesis H1 in these conditions, and likely for other, similar scenarios.
We shall discuss this in Chapter \ref{chap:discussion}.

Similarly, expressed in the same terms, H2 posits that 
our stacked user modeling should outperform the aggregation approaches
given in Table \ref{table:results:e1}.
The \emph{median} and \emph{average} aggregation methods
serve as globalized and generalized aggragation methods,
Stacked user modeling is adaptive in that each prediction is 
aggregated based on the current user and item,
satisfying the language of H2.

As we can see in Table \ref{table:results:e1:sum},
\emph{H2 is confirmed for these methods and this dataset}.
However, as our collection of aggregation methods is a lot simpler
than our collection of recommender systems, the strength of this combination
is notably weaker than that of H1.
Still, the fact that stacked user modeling outperforms these simple aggregation
approaches is a positive result warranting further experiments.
This will also be discussed in Chapter \ref{chap:discussion}.

It would seem then that, based on our experiments, available data
and assumptions of evaluation measures, both H1 and H2 are confirmed.
Our adaptive aggregation approach outperforms both standard recommender
methods and simple generalized aggregation methods.
Notably, our approach is more complex than the methods it outperforms,
so the question whether the methods performance is worth its extra complexity becomes important.
We shall discuss this, and other implications of these results in the next chapter.
For now, let us proceed to the second experiment and hypothesis H3.


